---
title: "postGGIR"
author: "Wei Guo"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{postGGIR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```



# Introduction


## What is GGIR?

[GGIR](https://CRAN.R-project.org/package=GGIR) is an R-package to process multi-day raw accelerometer data for physical activity and sleep research. GGIR will write all output files into two sub-directories of ./meta and ./results. GGIR is increasingly being used by a number of academic institutes across the world.  
 

## What is postGGIR?

[postGGIR](https://CRAN.R-project.org/package=postGGIR) is an R-package to data processing after running GGIR for accelerometer data. In detail, all necessary R/Rmd/shell files were generated for data processing after running GGIR for accelerometer data. Then in part 1, all csv files in the GGIR output directory were read, transformed and then merged. In part 2, the GGIR output files were checked and summarized in one excel sheet. In part 3, the merged data was cleaned according to the number of valid hours on each night and the number of valid days for each subject. In part 4, the cleaned activity data was imputed by the average ENMO over all the valid days for each subject. Finally, a comprehensive report of data processing was created using Rmarkdown, and the report includes few explortatory plots and multiple commonly used features extracted from minute level actigraphy data in part 5-7. This vignette provides a general introduction to postGGIR.

 


# Setting up your work environment

## Install R and RStudio

[Download and install R](https://cran.r-project.org/)

[Download and install RStudio](https://www.rstudio.com/products/rstudio) (optional, but recommended)

Download GGIR with its dependencies, you can do this with one command from the console command line:

```{R,eval=FALSE}
install.packages("postGGIR", dependencies = TRUE)
```

## Prepare folder structure

1. folder of .bin files for GGIR or a file listing all .bin files
    - R program will check the missing in the GGIR output by comparing with all raw .bin files
    
2. foder of the GGIR output with two sub-folders 
    - meta (./basic, ./csv, etc)     
    - results (part*summary*.csv)       
 


# Quick start


## Create a template shell script of postGGIR 


```{R,eval=FALSE}
library(postGGIR)
create.postGGIR()
```


The function will create a template shell script of postGGIR in the current directory, names as *STUDYNAME_part0.maincall.R*.
 


```{bash, eval=FALSE} 
cat STUDYNAME_part0.maincall.R
```


```{r,eval=FALSE}
options(width=2000) 
argv = commandArgs(TRUE);  
print(argv) 
print(paste("length=",length(argv),sep=""))  
mode<-as.numeric(argv[1])  
print(c("mode =", mode))
# (Note) Please remove the above lines if you are running this within R console 
#        instead of submitting jobs to a cluster.
 
#########################################################################   
# (user-define 1) you need to redefine this according different study!!!!
######################################################################### 
# example 1 
filename2id.1<-function(x)  unlist(strsplit(y1,"\\."))[1] 
 
#  example 2 (use csv file =c("filename","ggirID")) 
filename2id.2<-function(x) {
  d<-read.csv("./postGGIR/inst/example/filename2id.csv",head=1,stringsAsFactors=F)
  y1<-which(d[,"filename"]==x)
  if (length(y1)==0) stop(paste("Missing ",x," in filename2id.csv file",sep=""))
  if (length(y1)>=1) y2<-d[y1[1],"newID"] 
  return(as.character(y2))
} 


#########################################################################  
#  main call
######################################################################### 
  
call.afterggir<-function(mode,filename2id=filename2id.1){ 

library(postGGIR) 
#########################################################################  
# (user-define 2) Fill in parameters of your ggir output
########################################################################## 
  
currentdir = 
studyname =
bindir = 
outputdir =  
setwd(currentdir) 

rmDup=FALSE   # keep all subjects in postGGIR
PA.threshold=c(50,100,400)
part5FN="WW_L50M125V500_T5A5" 
epochIn = 5
epochOut = 5
flag.epochOut = 60
use.cluster = FALSE
log.multiplier = 9250
QCdays.alpha = 7
QChours.alpha = 16 
useIDs.FN<-NULL 
Rversion="R" 
desiredtz="US/Eastern" 
RemoveDaySleeper=FALSE 
part5FN=part5FN,
NfileEachBundle=20 
trace=FALSE
#########################################################################  
#   remove duplicate sample IDs for plotting and feature extraction 
######################################################################### 
if (mode==3 & rmDup){
# step 1: read ./summary/*remove_temp.csv file (output of mode=2)
keep.last<-TRUE #keep the latest visit for each sample
sumdir<-paste(currentdir,"/summary",sep="")  
setwd(sumdir)  
inFN<-paste(studyname,"_samples_remove_temp.csv",sep="")
useIDs.FN<-paste(sumdir,"/",studyname,"_samples_remove.csv",sep="") 

#########################################################################  
# (user-define 3 as rmDup=TRUE)  create useIDs.FN file
######################################################################### 
# step 2: create the ./summary/*remove.csv file manually or by R commands
d<-read.csv(inFN,head=1,stringsAsFactors=F)
d<-d[order(d[,"Date"]),]
d<-d[order(d[,"newID"]),]
d[which(is.na(d[,"newID"])),]
S<-duplicated(d[,"newID"],fromLast=keep.last) #keep the last copy for nccr
d[S,"duplicate"]<-"remove"
write.csv(d,file=useIDs.FN,row.names=F)

} 

#########################################################################  
#   call afterggir
######################################################################### 

setwd(currentdir)  
afterggir(mode=mode,
          useIDs.FN=useIDs.FN,
          currentdir=currentdir,
          studyname=studyname,
          bindir=bindir,
          outputdir=outputdir,
          epochIn=epochIn,
          epochOut=epochOut,
          flag.epochOut=flag.epochOut,
          log.multiplier=log.multiplier,
          use.cluster=use.cluster,
          QCdays.alpha=QCdays.alpha,
          QChours.alpha=QChours.alpha,
          QCnights.feature.alpha=QCnights.feature.alpha, 
          Rversion=Rversion,
          filename2id=filename2id,
          PA.threshold=PA.threshold,
          desiredtz=desiredtz,
          RemoveDaySleeper=RemoveDaySleeper,
          part5FN=part5FN,
          NfileEachBundle=NfileEachBundle,
          trace=trace) 

} 
#########################################################################
          call.afterggir(mode)   
######################################################################### 

#   Note:   call.afterggir(mode)
#        mode = 0 : creat sw/Rmd file
#        mode = 1 : data transform using cluster or not
#        mode = 2 : summary
#        mode = 3 : clean 
#        mode = 4 : impu
```

## Edit shell script 

Three places were marked as "user-define" and need to be edited by user in the *STUDYNAME_part0.maincall.R* file. Please rename the file by replacing your real studyname after the edition.

### 1. Define the function filename2id( )    
This user-defined function will change the filename of the raw accelerometer file to the short ID. For example, the first example change "0002__026907_2016-03-11 13-05-59.bin" to new ID of "0002". If you prefer to define new ID by other way, you could create a .CSV file including "filename" and "newID" at least and then defined this function as the second example. The new variable of "newID", included in the output files, could be used as the key ID in the summary report of postGGIR and be used to define the duplicate samples as well. 


### 2. Parameters of shell script  
User needs to define the following parameters as follows,
 
 

Variables	| Description
----------------- | ----------------------------------------------------
rmDup | Set rmDup = TRUE if user want to remove some samples such as duplicates. Set rmDup = FALSE if user want to keep all samples.    
mode |   Specify which of the five parts need to be run, e.g. mode = 0 makes that all R/Rmd/sh files are generated for other parts. When mode = 1, all csv files in the GGIR output directory were read, transformed and then merged. When mode = 2, the GGIR output files were checked and summarized in one excel sheet. When mode = 3, the merged data was cleaned according to the number of valid hours on each night and the number of valid days for each subject. When mode = 4, the cleaned data was imputed. 
useIDs.FN |  Filename with or without directory for sample information in CSV format, which including "filename" and "duplicate" in the headlines at least. If duplicate="remove",  the accelerometer files will not be used in the data analysis of part 5-7. Defaut is NULL, which makes all accelerometer files will be used in part 5-7.
currentdir  |  Directory where the output needs to be stored. Note that this directory must exist.  
studyname  |   Specify the study name that used in the output file names
bindir  |   Directory where the accelerometer files are stored or list   
outputdir  |  Directory where the GGIR output was stored.  
epochIn  |  Epoch size to which acceleration was averaged (seconds) in GGIR output. Defaut is 5 seconds.
epochOut  |  Epoch size to which acceleration was averaged (seconds) in part1. Defaut is 5 seconds.
flag.epochOut  |    Epoch size to which acceleration was averaged (seconds) in part 3. Defaut is 60 seconds.  
log.multiplier  | The coefficient used in the log transformation of the ENMO data, i.e. log( log.multiplier * ENMO + 1), which have been used in part 5-7. Defaut is 9250. 
use.cluster  |   Specify if part1 will be done by parallel computing. Default is TRUE, and the CSV file in GGIR output will be merged for every 20 files first, and then combined for all. 
QCdays.alpha  |    Minimum required number of valid days in subject specific analysis as a quality control step in part2. Default is 7 days. 
QChours.alpha  |  Minimum required number of valid hours in day specific analysis as a quality control step in part2. Default is 16 hours. 
QCnights.feature.alpha  |   Minimum required number of valid nights in day specific mean and SD analysis as a quality control step in the JIVE analysis. Default is  c(0,0), i.e. no additional data cleaning in this step.  
Rversion |   R version, eg. "R/3.6.3". Default is "R". 
filename2id  |   User defined function for converting filename to sample IDs. Default is NULL.  
PA.threshold  |  Threshold for light, moderate and vigorous physical activity. Default is c(50,100,400).
desiredtz |   desired timezone: see also http://en.wikipedia.org/wiki/Zone.tab. Used in g.inspectfile(). Default is "US/Eastern". 
RemoveDaySleeper  |   Specify if the daysleeper nights are removed from the calculation of number of valid days for each subject. Default is FALSE. 
part5FN   |    Specify which output is used in the GGIR part5 results. Defaut is "WW_L50M125V500_T5A5", which means that part5_daysummary_WW_L50M125V500_T5A5.csv and part5_personsummary_WW_L50M125V500_T5A5.csv are used in the analysis. 
NfileEachBundle  |   Number of files in each bundle when the csv data were read and processed in a cluster. Default is 20. 
trace  |   Specify if the intermediate results is printed when the function was executed. Default is FALSE.
 





### 3. Subset of samples (optional)    
The postGGIR package not only simply transform/merge the activity and sleep data, but it also can do some prelimary data analysis such as principle componet analysis and feature extraction. Therefore, the basic data clean will be processed first as follows,     

   - data clean by removing valid days/samples defined by minimum required number of valid hours/days in the activity data   
   - remove duplicate samples  

If you prefer to use all samples, just skip this part and use `rmDup=FALSE` as the default. Otherwise, if you want to remove some samples such as duplicates, there are two ways as follows,

   -  Edit R codes of "step 2" in this part. For example, the template will keep the later copy for duplicate samples    
   -  Remove R codes of "step 2" in this part, and create *studyname_samples_remove.csv* file by filling "remove" in the "duplicate" column in the template file of *studyname_samples_remove_temp.csv*. The data will be kept unless duplicate="remove".  



## Run R script 
 

```{R,eval=FALSE} 
call.afterggir(mode,filename2id)   
```



Variables	| Description
----------------- | ----------------------------------------------------
mode | Specify which of the five parts need to be run, e.g. mode = 0 makes that all R/Rmd/sh files are generated for other parts. When mode = 1, all csv files in the GGIR output directory were read, transformed and then merged. When mode = 2, the GGIR output files were checked and summarized in one excel sheet. When mode = 3, the merged data was cleaned according to the number of valid hours on each night and the number of valid days for each subject. When mode = 4, the cleaned data was imputed.  
filename2id | This user-defined function will change the filename of the raw accelerometer file to the short ID for the purpose of identifying duplicate IDs. 



## Run script in a cluster

```{R,eval=FALSE} 
#!/bin/bash
#
#$ -cwd
#$ -j y
#$ -S /bin/bash
  source ~/.bash_profile

   cd /postGGIR/inst/example/afterGGIR; 
   module load R ; 
     R --no-save --no-restore --args  < studyname_ggir9s_postGGIR.pipeline.maincall.R  0
     R --no-save --no-restore --args  < studyname_ggir9s_postGGIR.pipeline.maincall.R  1
     R --no-save --no-restore --args  < studyname_ggir9s_postGGIR.pipeline.maincall.R  2
     R --no-save --no-restore --args  < studyname_ggir9s_postGGIR.pipeline.maincall.R  3
     R --no-save --no-restore --args  < studyname_ggir9s_postGGIR.pipeline.maincall.R  4 

     R -e "rmarkdown::render('part5_studyname_postGGIR.report.Rmd'   )" 
     R -e "rmarkdown::render('part6_studyname_postGGIR.nonwear.report.Rmd'   )" 
     R -e "rmarkdown::render('part7a_studyname_postGGIR_JIVE_1_somefeatures.Rmd'   )" 
     R -e "rmarkdown::render('part7b_studyname_postGGIR_JIVE_2_allfeatures.Rmd'   )" 
     R -e "rmarkdown::render('part7c_studyname_postGGIR_JIVE_3_excelReport.Rmd'   )" 

```

# Software Functionalities  
## Part 1  
The functions of part 1 read the activity data measured by Euclidian norm minus one (ENMO) in long csv-spreadsheets and then the data was merged and transformed into a square matrix for all subjects. To be noted, a day could have 23 or 25 hours when daylight saving time started or ended. The ENMO data would take the average for those duplicate timestamps caused by daylight saving time so as to be merged systematically with other regular 24-hour days.  In addition, the angle of the z-axis (ANGLEZ) relative to the horizontal plane (degrees) utilizing all three axes was also merged and saved into an excel file. Besides that, more data were output including the light mean, light peak, temperature mean and clipping score and Euclidian norm metric (EN) if available. All outputs were stored under the ./data directory for part 1.  

## Part 2  
The functions of part 2 introduced some descriptive variables of all accelerometer files in the GGIR output. In part 2, an excel file was output under the ./summary directory, which includes nine pages as follows, 1) List of files in the GGIR output (2) Summary of numbers of output files (3) List of duplicate IDs (4) ID errors (5) Number of valid days (6) Table of number of valid/missing days (7) Missing pattern (8) Frequency of the missing pattern (9) Description of all accelerometer files.  Multiple plots were generated in a pdf file including the number of valid hours, days, missing pattern, etc.  Technically, the raw accelerometer data was visited to obtain a complete input list for the purpose of examining missingness of GGIR output when user specified the path. Additionally, the summary results from GGIR output were also visited here to form a comprehensive data quality report in the part 2 of postGGIR. 
    
## Part 3 and Part 4  
The functions of part 3 introduced some 'flag' variables for data cleaning of the merged ENMO and ANGLEZ data. As default, only days with more than 16 hours were marked as valid days and subjects with more than 7 valid days were marked as valid samples. User could specify these two parameters in the main function. Further, the data could be aggregated into minute-level or hour-level data as required by users. In part 4, the ENMO data was imputed by taking the subject-level mean over all the valid days for each subject. All output was stored under the ./data directory. In the output file, a few description variables were included as follows (1) the number of valid hours; (2) missing pattern for each subject; (3) non-wear time in minutes; (4) an indicator variable to indicate if the visit should be removed for having multiple visits and (5) the number of missing values after imputation. Samples were removed when systematic missingness was observed, i.e., the activity data was missing on same timestamps among all days and therefore could not be imputed.  

## Part 5  
The functions of part 5 generated a comprehensive report in the .html format, which included data quality check and an exploratory data analysis. In the report, first, the file lists were summarized in each GGIR folders and summary files so that user could easily check the missingness in each GGIR parts. Second, the duplicate samples were checked and marked, where the duplication might be caused by having multiply visits for some samples but only one visit will be kept in the data analysis such as functional principal component analysis (FPCA). Thirdly, as shown in Figure 1, data quality was checked such as the number of valid days, non-wear time and missing pattern, which was helpful to better understand the data by the graphical presentation. As an exploratory analysis, the data correlation, and the output of FPCA analysis was plotted in the report.

## Part 6 (Optional) 
The non-wear data was loaded from the M$metalong$nonwearscore variable of the R data that was stored in the folder of /meta/basic/meta of the GGIR output, and the matrix to clarify when data was imputed for each long epoch time window and the reason for imputation.  This function will generate a non-wear matrix at minute level, and it could be skipped if user chose to use the imputation data in the JIVE application as default.
 
## Part 7  
In part 7, about 100 features were extracted from minute level activity data of three domains of sleep, physical activity, and circadian rhythmicity, which were based on outputs from GGIR v2.4.0 and calculated by R ActFrag and ActCR packages (Di et al., 2019). The standard deviation across days on each subject was also created for each feature. The weekday and weekend specific features were extracted as well since most features in the sleep and physical activity showed significant difference between weekdays and weekends. In brief, sleep domain referred to sleep duration, midpoint, efficiency, etc. Physical activity referred to daily motor activity such as sedentary behavior, light, and moderate-to-vigorous physical activity (MVPA). Circadian rhythms were natural rhythms that regulates the sleep-wake cycle within every 24 hours. For example, the cosinor curve and FPCA analysis were used in modeling of biological rhythms. A comprehensive list of all features could be found in the supplementary table and the detailed definition could be found in the GGIR manual and Di et al.'s publication in 2019.
 
	
	 
	

# Running postGGIR and Inspecting the results
 
## Input and output of part 0

  *  Command = call.afterggir(mode=0)    
  *  Output folder = ./

Output         |    Description
-------------- | ------------------------------------------------ 
part1_data.transform.R (use.cluster=TRUE, optional) | R code for data transformation and merge for every 20 files in each partition. When the number of .bin files is large ( > 1000), the data merge could take long time, user could split the job and submit the job to a cluster for parallel computing.
part1_data.transform.sw  (use.cluster=TRUE, optional) | Submit the job to a cluster for parallel computing
part1_data.transform.merge.sw  (use.cluster=TRUE, optional) | Merge all partitions for the ENMO and ANGLEZ data  
part5_studyname_postGGIR.report.Rmd | 	R markdown file for generate  a comprehensive report of data processing  and explortatory plots.
part6_studyname_postGGIR.nonwear.report.Rmd | 	R markdown file for generate  a report of nonwear score.
part7a_studyname_postGGIR_JIVE_1_somefeatures.Rmd | 	Extract some features from the actigraphy data using R
part7b_studyname_postGGIR_JIVE_2_allfeatures.Rmd | 	Extract other features from the GGIR output and merge all features together 
part7c_studyname_postGGIR_JIVE_3_runJIVE.Rmd	| Perform JIVE Decomposition for All Features using r.jive
part7d_studyname_postGGIR_JIVE_4_somefeatures_weekday.Rmd	| Extract some weekday/weekend specific features from the actigraphy data using R

part9_swarm.sh | shell script to submit all jobs to the cluster


## Input and output of part 1
  +  Main input files: csv files under /meta/csv folders of GGIR output 
  +  Command = call.afterggir(mode=1) 
  +  Output folder = ./data
 
Output | Description
-------------- | --------------------------------------- 
studyname_filesummary_csvlist.csv | File list in the ./csv folder of GGIR
studyname_filesummary_Rdatalist.csv | File list in the ./basic folder of GGIR	
All_studyname_ANGLEZ.data.csv | Raw data of ANGLEZ after merge
All_studyname_ENMO.data.csv | Raw data of ENMO after merge	
nonwearscore_studyname_f0_f1_Xs.csv | Data matrix of nonwearscore
nonwearscore_studyname_f0_f1_Xs.pdf | Plots for nonwearscore
plot.nonwearVSnvalidhours.csv | Nonwear data for plot
plot.nonwearVSnvalidhours.pdf | Nonwear plots	
lightmean_studyname_f0_f1_Xs.csv | Data matrix of lightmean
lightpeak_studyname_f0_f1_Xs.csv | Data matrix of lightpeak
temperaturemean_studyname_f0_f1_Xs.csv | Data matrix of temperaturemean
clippingscore_studyname_f0_f1_Xs.csv | Raw data of clippingscore
EN_studyname_f0_f1_Xs.csv | Data matrix of EN

 

*f0 and f1 are the file index to start and finish with  
*Xs is the epoch size to which acceleration was averaged (seconds) in GGIR output


## Input and output of part 2
  + Main input files     
         +    ./data/All_studyname_ENMO.data.csv    
         +     GGIR results: part2, part4 and part5 (please specify $part5FN$ in the main function)   
         +     GGIR raw data when bindir was specified   

  +  Command = call.afterggir(mode=2)  
  +  Output folder = ./summary
 
Output | Description
-------------- | ----------------------------------------- 
studyname_ggir_output_summary.xlsx | Description of all accelerometer files in the GGIR output.  This excel file includs 9 pages as follows,  (1) List of files in the GGIR output (2) Summary of files (3) List of duplicate IDs  (4)  ID errors (5) Number of valid days (6) Table of number of valid/missing days (7) Missing patten (8) Frequency of the missing pattern  (9) Description of all accelerometer files.
part2daysummary.info.csv | Intermediate results for description of each accelerometer file.
studyname_ggir_output_summary_plot.pdf | Some plots such as the number of valid days, which were included in the part2a_studyname_postGGIR.report.html file as well.
studyname_samples_remove_temp.csv | Create studyname_samples_remove.csv file by filling "remove" in the "duplicate" column in this template. If duplicate="remove",  the accelerometer files will not be used in the data analysis of part5.



## Input and output of part 3
  +  Main input file: ./data/All_studyname_ENMO.data.csv  
  +  Command = call.afterggir(mode=3) 
  +  Output folder = ./data
 
Output | Description
-------------- | -------------------------------------- 
flag_All_studyname_ANGLEZ.data.Xs.csv | Adding flags for data cleaning of the raw ANGLEZ data
flag_All_studyname_ENMO.data.Xs.csv | Adding flags for data cleaning of the raw ENMO data
IDMatrix.flag_All_studyname_ENMO.data.60s.csv | ID matrix



*Xs is the epoch size to which acceleration was averaged (seconds) in GGIR output



## Input and output of part 4
  +  Main input file: ./data/flag_All_ studyname _ENMO.data.5s.csv 
  +  Command = call.afterggir(mode=4) 
  +  Output folder = ./data
 
Output | Description
-------------- | -------------------------------------------------------------------------------- 
impu.flag_All_studyname_ENMO.data.60s.csv | Imputation data for the merged ENMO data, and the missing values were imputated by the average ENMO over all the valid days for each subject. 
	


## Description of flag variables in the output data

Variable | Description
-------------- | --------------------------------------------------------------------------------
filename | accelerometer file name 
Date | date recored from the GGIR part2.summary file 
id | IDs recored from the GGIR part2.summary file 
calender_date | date in the format of yyyy-mm-dd 
N.valid.hours | number of hours with valid data recored from the part2_daysummary.csv file in the GGIR output 
N.hours | number of hours of measurement recored from the part2_daysummary.csv file in the GGIR output 
weekday | day of the week-Day of the week 
measurementday | day of measurement-Day number relative to start of the measurement 
newID | new IDs defined as the user-defined function of filename2id(), e.g. substrings of the filename 
Nmiss_c9_c31 | number of NAs from the 9th to 31th column in the part2_daysummary.csv file in the GGIR output
missing | "M" indicates missing for an invalid day, and  "C" indicates completeness for a valid day 
Ndays | number of days of measurement  
ith_day | rank of the measurementday, for example, the value is 1,2,3,4,-3,-2,-1 for measurementday = 1,...,7 
Nmiss | number of missing (invalid) days  
Nnonmiss | number of non-missing (valid) days 
misspattern | indicators of missing/nonmissing for all measurement days at the subject level 
RowNonWear | number of columnns in the non-wearing matrix 
NonWearMin | number of minutes of non-wearing
remove16h7day | indicator of a key qulity control output. If remove16h7day=1, the day need to be removed. If remove16h7day=0, the day need to be kept.
duplicate | If duplicate="remove",  the accelerometer files will not be used in the data analysis of part5.
ImpuMiss.b | number of missing values on the ENMO data before imputation
ImpuMiss.a | number of missing values on the ENMO data after imputation
KEEP | The value is "keep"/"remove", e.g. KEEP="remove" if remove16h7day=1 or duplicate="remove" or ImpuMiss.a>0



## Input and output of part 5
  + Main input files
      +  ./summary/studyname_ggir_output_summary.xlsx
      +  ./summary/part24daysummary.info.csv
      +  ./data/plot.nonwearVSnvalidhours.csv
      +  ./data/impu.flag_All_studyname_ENMO.data.flag.epochOuts.csv 

  +  Command: run part5_studyname_postGGIR.report.Rmd 
  +  Output folder = ./ 
 
Output | Description
-------------- | --------------------------------------------------------------------------------  
part5_studyname_postGGIR.report.html | A comprehensive report of data processing  and explortatory plots.


## Input and output of part 6
  +  Main input file: nonwearscore_studyname_01_xx_900s.csv 
  +  Command = run part6_studyname_postGGIR.nonwear.report.Rmd
  +  Output folder = ./ 

Folder | Output | Description
-------------- | -------------- | -------------------------------------------------------------------------------- 
./ | part6_studyname_postGGIR.nonwear.report.html | A report of nonwear score.
./data | JIVEraw_nonwearscore_studyname_f0_f1_Xs.csv | Imputation data matrix of nonwearscore (1/0)
./data | JIVEimpu_nonwearscore_studyname_f0_f1_Xs.csv | Data matrix of nonwearscore (1/0/NA)



*f0 and f1 are the file index to start and finish with  
*Xs is the epoch size to which acceleration was averaged (seconds) in GGIR output


## Input and output of part 7a

  +  Main input 
       +  ./data/impu.flag_All_studyname_ENMO.data.flag.epochOuts.csv
       +  GGIR: /results/QC/part4_nightsummary_sleep_full.csv 

  +  Command = run part7a_studyname_postGGIR_JIVE_1_somefeatures.Rmd
  +  Output folder = ./ 

Output | Description
-------------- | --------------------------------------------------------------------------------
part7_studyname_all_features_dictionary.xlsx | Description of features
part7a_studyname_postGGIR_JIVE_1_somefeatures.html | Extract some features from the actigraphy data using R
part7a_studyname_some_features_page1_features.csv | List of some features
part7a_studyname_some_features_page2_face_day_PCs.csv | Function PCA at the day level using fpca.face( )
part7a_studyname_some_features_page3_face_subject_PCs.csv | Function PCA at the subject level using fpca.face( )
part7a_studyname_some_features_page4_denseFLMM_day_PCs.csv | Function PCA at the day level using denseFLMM( )
part7a_studyname_some_features_page5_denseFLMM_subject_PCs.csv | Function PCA at the subject level using denseFLMM( )


## Input and output of part 7b
  +  Main inputs
     +  GGIR: part2_summary.csv 
     +  GGIR:  part2_daysummary.csv
     +  GGIR: part4_nightsummary_sleep_cleaned.csv 
     +  GGIR: /results/QC/part4_nightsummary_sleep_full.csv 
     +  GGIR:  part5_daysummary_part5FN.csv  
     +  part7_studyname_all_features_dictionary.xlsx
 
  +  Command = run part7b_studyname_postGGIR_JIVE_2_allfeatures.Rmd
  +  Output folder = ./ 


Output | Description
-------------- | --------------------------------------------------------------------------------
part7b_studyname_postGGIR_JIVE_2_allfeatures.html | Extract other features from the GGIR output and merge all features together
part7b_studyname_all_features_1.csv | Raw data of all features  
part7b_studyname_all_features_2.csv | Keep sample with valid ENMO inputs
part7b_studyname_all_features_2.csv.log | Log file of each variable of part5b_studyname_all_features_2.csv
plot_part7b_studyname_all_features_2.csv.pdf | Plot of each variable of part5b_studyname_all_features_2.csv
part7b_studyname_all_features_3.csv | Average variable at the subject level
part7b_studyname_all_features_3.csv.log | Log file of each variable of part5b_studyname_all_features_3.csv
plot_part7b_studyname_all_features_3.csv.pdf | Plot of each variable of part5b_studyname_all_features_3.csv
part7b_studyname_all_features_4.csv | subject level SD of each feature

 

## Input and output of part 7c
  + Main inputs 
     +  part7b_studyname_all_features_3_subject.csv
     +  part7b_studyname_all_features_4_subjectSD.csv

  +  Command = run part7c_studyname_postGGIR_JIVE_3_runJIVE.Rmd 
  +  Output folder = ./  

Output | Description
-------------- | --------------------------------------------------------------------------------
part7c_studyname_postGGIR_JIVE_4_outputReport.html | Perform JIVE Decomposition for All Features using r.jive
part7c_studyname_jive_Decomposition.csv | Joint and individual structure estimates
part7c_studyname_jive_predScore.csv | PCA scores of JIVE ( missing when jive.predict failes)
part7c_studyname_jive_predScore.csv | PCA scores of JIVE ( missing when jive.predict failes)



 


## Input and output of part 7d
  +  Main input: ./data/impu.flag_All_studyname_ENMO.data.flag.epochOuts.csv 
  +  Command = run part7d_studyname_postGGIR_JIVE_5_somefeatures_weekday.Rmd
  +  Output folder = ./  

Output | Description
-------------- | ------------------------------------- 
part7d_studyname_some_features_page1.csv | Perform JIVE Decomposition for All Features using r.jive
part7d_weekday_studyname_all_features_3.csv | subject level mean of each feature on weekday
part7d_weekday_studyname_some_features_page4_denseFLMM_day_PCs.csv | Function PCA at the day level using denseFLMM( ) on weekday
part7d_weekday_studyname_some_features_page5_denseFLMM_subject_PCs.csv | Function PCA at the subject level using denseFLMM( ) on weekday
part7d_weekend_studyname_all_features_3.csv | subject level mean of each feature on weekend
part7d_weekend_studyname_some_features_page4_denseFLMM_day_PCs.csv | Function PCA at the day level using denseFLMM( ) on weekend
part7d_weekend_studyname_some_features_page5_denseFLMM_subject_PCs.csv | Function PCA at the subject level using denseFLMM( ) on weekend





## Description of features of domains of physical activity, sleep and circadian rhythmicity

**Sleep Domain**  


```{r,echo=F}
library(xlsx)   
library(knitr)  
library(kableExtra)  


#feaFN<-system.file("template", "features.dictionary.xlsx", package = "postGGIR")  
feaFN<- "/data/guow4/project0/GGIR/postGGIR/postGGIR_compile/v2/postGGIR/inst/template/features.dictionary.xlsx"   
  
dict<-read.xlsx(feaFN,head=1,sheetName="dictionary",stringsAsFactors=F)
dict.SL<-dict[which(dict[,"Domain"]=="SL"),c("Variable","Description")]
dict.PA<-dict[which(dict[,"Domain"]=="PA"),c("Variable","Description")] 
dict.CR<-dict[which(dict[,"Domain"]=="CR"),c("Variable","Description")]

row.names(dict.SL)<-NULL
row.names(dict.PA)<-NULL
row.names(dict.CR)<-NULL

kable(dict.SL) %>%
     kable_styling(bootstrap_options = c("striped", "hover"))
```
 


**Physical Activity Domain**      
 
```{r,echo=F}
kable(dict.PA) %>%
     kable_styling(bootstrap_options = c("striped", "hover"))
```

**Circadian Rhythmicity Domain**    
  
```{r,echo=F}
kable(dict.CR) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
```


```{r,eval=FALSE,include=FALSE} 
d1<-read.xlsx("postGGIR.output.description.xlsx",sheetName="output.format")  

cd /data/guow4/project0/GGIR/postGGIR/postGGIR_compile/v2/postGGIR/vignettes     
R -e "rmarkdown::render('postGGIR.Rmd'   )"  
```


#
# Reference 
    
 +  Wei Guo,  Vadim Zipunnikov, Andrew Leroux, PhD, Kathleen R Merikangas (2021) postGGIR: An Open-Source R/R-Markdown Package for Accelerometer Data Processing after Running GGIR. (Manuscript)  
  